An elaborate data set on human gait and the effect of mechanical perturbations

### 2014-12-04 paper.tex
Added a title: closes #32.
Introduced the F1000 sty file.

\documentclass{article}

\usepackage{graphicx}
\usepackage{booktabs} % for table /toprule, /midrule, etc
\usepackage{todo}

\title{Perturbed Gait Data}

@@

@@

\end{abstract}

\section{Introduction}
%
Emerging powered prosthetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an able-bodied person. A modern gait lab is able to
collect a variety of kinematic, kinetic, and physiological data from humans
during gait. This data can potentially be used to drive the design of the human
mimicking controller. With a rich enough data set, one may be able to identify
control mechanism used during a human's natural gait and recovery from
perturbations. We have collected data that we believe is richer than previous
gait data sets and may be rich enough for control identification. The data can
also be used for verification purposes for controllers that have been designed
in other manners.

@@

\section{Equipment}
%
The Human Motion and Control Lab at Cleveland State University has modern human
motion measurement equipment. Our system includes:

@@

\section{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the ``mocap'' and ``record'' modules in D-Flow in addition to a YAML file
that contains all of the necessary meta data for the given trial. These three
files are stored in a hierarchy of directories with one trial per directory.
The directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

@@

\subsection{mocap-xxx.txt}

@@

\subsection{record-xxx.txt}
%
The record module also outputs a tab delimited ASCII text file. It includes a
\verb+Time+ column which records the D-Flow system time in seconds which
corresponds to the time recorded in the \verb+TimeStamp+ column in mocap module
tsv file. There are two additional columns \verb+RightBeltSpeed+ and
\verb+LeftBeltSpeed+ which provide the independent belt speeds measured in
meters per second by an encoder in the treadmill.

@@

\subsection{meta-xxx.yml}

@@

\begin{figure}
  \small
  \begin{verbatim}
    study:
            id: 1
            name: Gait Control Identification
            description: Perturb the subject during walking and running.
    subject:
            id: 0
    trial:
            id: 58
            subject-id: 0
            datetime: 2014-03-28
            notes: >
                    This is an unloaded trial (no subject) 10 minute protocol that
                    perturbs the subject longitudinally. The treadmill reference
                    markers were not in place. We have the four custom made metal
                    feet placed under the four corners of the treadmill to prevent
                    pitch and sway movements. In addition, we also have the front
                    left and rear right feet that were installed by Motek. We also
                    have +/-3g analog accelerometers (ADXL330) affixed to the four
                    corners of the main frame of the treadmill and wired into the
                    NI box. The right handrail is removed.
            nominal-speed: 0.8
            nominal-speed-units: meters per second
            stationary-platform: True
            pitch: False
            sway: False
            hardware-settings:
                    high-performance: True
            analog-channel-map:
                    Channel1.Anlg: F1Y1
                    Channel2.Anlg: F1Y2
                    Channel3.Anlg: F1Y3
                    Channel4.Anlg: F1X1
                    Channel5.Anlg: F1X2
                    Channel6.Anlg: F1Z1
                    Channel7.Anlg: F2Y1
                    Channel8.Anlg: F2Y2
                    Channel9.Anlg: F2Y3
                    Channel10.Anlg: F2X1
                    Channel11.Anlg: F2X2
                    Channel12.Anlg: F2Z1
                    Channel61.Anlg: Front_Left_AccX
                    Channel62.Anlg: Front_Left_AccY
                    Channel63.Anlg: Front_Left_AccZ
                    Channel64.Anlg: Back_Left_AccX
                    Channel65.Anlg: Back_Left_AccY
                    Channel66.Anlg: Back_Left_AccZ
                    Channel67.Anlg: Front_Right_AccX
                    Channel68.Anlg: Front_Right_AccY
                    Channel69.Anlg: Front_Right_AccZ
                    Channel70.Anlg: Back_Right_AccX
                    Channel71.Anlg: Back_Right_AccY
                    Channel72.Anlg: Back_Right_AccZ
            events:
                    A: Force Plate Zeroing
                    B: Calibration Pose
                    C: First Normal Walking
                    D: Longitudinal Perturbation
                    E: Second Normal Walking
                    F: Unloaded End
            files:
                    mocap: mocap-058.txt
                    record: record-058.txt
                    meta: meta-058.yml
  \end{verbatim}
  \caption{An example meta data YAML file.}
  \label{fig:example-meta-data}
\end{figure}

@@

\subsection{Markers}
%
We make use of the full body 47 marker set described in \cite{Ton's HBM paper}
and presented in detail in Table \ref{tab:marker-labels}. As with all camera
based motion capture systems, the markers sometimes go missing in the
recording.

@@

\begin{table}

@@

\end{table}

\section{Perturbation Signals}
%
The data set uses a testing protocol containing a segment of normal walking,
followed by a series of pseudo-random longitudinal perturbations, and ending
with a second segment of normal walking. Three pseudo-random speed control
signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
generated through MATLAB and Simulink. Initially, a  random acceleration signal
was created using discrete-time Gaussian white noise and saturated at the maximum
belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt speed, and 
high-pass filtered (second-order Butterworth) to eliminate velocity drift. Mean 
velocity was then added to the signal and limited between 0 and 3.6 m/s. The 
cutoff frequencies of the high-pass filter, as well as the variance in the acceleration 
signal, were adjusted until the appropriate standard deviations for each mean 
velocity were obtained.

@@

\begin{figure}
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt velocity input signals (purple) and recorded output 
                velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
                respectively}
  \label{fig:input_output}
\end{figure}

@@

\begin{figure}
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency analysis of the treadmill belt velocity input signal 
                 (purple) and the recorded output velocity (blue)
                for average belt speeds of 0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure}

\section{Processed Data}

@@

\begin{figure}
  \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
  \caption{Mean (solid, dashed) and $2\sigma$ (shaded) ($N=120$) joint angles
    and torques from both unperturbed (purple) and perturbed (blue) gait cycles
    from trial 20.}
  \label{fig:angle-torque-comparison}
\end{figure}

@@

\begin{figure}
  \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
  \caption{Box plots of the average belt speed, stride frequency, and stride
    length which compare 120 gait cycles for the unperturbed (purple) and
    perturbed (blue).}
  \label{fig:gait-cycle-stats-comparison}
\end{figure}

\section{Conclusion}

@@

\section{Acknowledgments}
We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.  The work was partially funded by the
State of Ohio Third Frontier Commission through the Wright Center for Sensor Systems Engineering (WCSSE).

>>>>

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage{f1000_styles}

\usepackage{graphicx}
\usepackage{booktabs} % for table /toprule, /midrule, etc

% I get a conflict with the todo stuff in f1000_styles if this is loaded.
%\usepackage{todo}

@@

\title{An elaborate data set on human gait and the effect of mechanical
perturbations}

\author[1]{Jason K. Moore}
\author[1]{Sandra K. Hnat}
\author[1]{Antonie van den Bogert}

\affil[1]{Mechanical Engineering, Cleveland State University, Cleveland, Ohio,
  USA, 44115. j.k.moore19@csuohio.edu, s.hnat@vikes.csuohio.edu,
  a.vandenbogert@csuohio.edu}

\maketitle
\thispagestyle{fancy}

@@

\end{abstract}
\listoftodos[F1000Research review comments] % Ignore until review stage
\clearpage

\section*{Introduction}
%
Emerging powered prosthetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an able-bodied person. A modern gait lab is able to
collect a variety of kinematic, kinetic, and physiological data from humans
during gait. This data can potentially be used to drive the design of the human
mimicking controller. With a rich enough data set, one may be able to identify
control mechanism used during a human's natural gait and recovery from
perturbations. We have collected data that we believe is richer than previous
gait data sets and may be rich enough for control identification. The data can
also be used for verification purposes for controllers that have been designed
in other manners.

@@

\section*{Equipment}
%
The Human Motion and Control Lab at Cleveland State University has modern human
motion measurement equipment. Our system includes:

@@

\section*{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the ``mocap'' and ``record'' modules in D-Flow in addition to a YAML file
that contains all of the necessary meta data for the given trial. These three
files are stored in a hierarchy of directories with one trial per directory.
The directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

@@

\subsection*{mocap-xxx.txt}

@@

\subsection*{record-xxx.txt}
%
The record module also outputs a tab delimited ASCII text file. It includes a
\verb+Time+ column which records the D-Flow system time in seconds which
corresponds to the time recorded in the \verb+TimeStamp+ column in mocap module
tsv file. There are two additional columns \verb+RightBeltSpeed+ and
\verb+LeftBeltSpeed+ which provide the independent belt speeds measured in
meters per second by an encoder in the treadmill.

@@

\subsection*{meta-xxx.yml}

@@

\begin{figure*}
  \small
  \begin{verbatim}
    study:
            id: 1
            name: Gait Control Identification
            description: Perturb the subject during walking and running.
    subject:
            id: 0
    trial:
            id: 58
            subject-id: 0
            datetime: 2014-03-28
            notes: >
                    This is an unloaded trial (no subject) 10 minute protocol that
                    perturbs the subject longitudinally. The treadmill reference
                    markers were not in place. We have the four custom made metal
                    feet placed under the four corners of the treadmill to prevent
                    pitch and sway movements. In addition, we also have the front
                    left and rear right feet that were installed by Motek. We also
                    have +/-3g analog accelerometers (ADXL330) affixed to the four
                    corners of the main frame of the treadmill and wired into the
                    NI box. The right handrail is removed.
            nominal-speed: 0.8
            nominal-speed-units: meters per second
            stationary-platform: True
            pitch: False
            sway: False
            hardware-settings:
                    high-performance: True
            analog-channel-map:
                    Channel1.Anlg: F1Y1
                    Channel2.Anlg: F1Y2
                    Channel3.Anlg: F1Y3
                    Channel4.Anlg: F1X1
                    Channel5.Anlg: F1X2
                    Channel6.Anlg: F1Z1
                    Channel7.Anlg: F2Y1
                    Channel8.Anlg: F2Y2
                    Channel9.Anlg: F2Y3
                    Channel10.Anlg: F2X1
                    Channel11.Anlg: F2X2
                    Channel12.Anlg: F2Z1
                    Channel61.Anlg: Front_Left_AccX
                    Channel62.Anlg: Front_Left_AccY
                    Channel63.Anlg: Front_Left_AccZ
                    Channel64.Anlg: Back_Left_AccX
                    Channel65.Anlg: Back_Left_AccY
                    Channel66.Anlg: Back_Left_AccZ
                    Channel67.Anlg: Front_Right_AccX
                    Channel68.Anlg: Front_Right_AccY
                    Channel69.Anlg: Front_Right_AccZ
                    Channel70.Anlg: Back_Right_AccX
                    Channel71.Anlg: Back_Right_AccY
                    Channel72.Anlg: Back_Right_AccZ
            events:
                    A: Force Plate Zeroing
                    B: Calibration Pose
                    C: First Normal Walking
                    D: Longitudinal Perturbation
                    E: Second Normal Walking
                    F: Unloaded End
            files:
                    mocap: mocap-058.txt
                    record: record-058.txt
                    meta: meta-058.yml
  \end{verbatim}
  \caption{An example meta data YAML file.}
  \label{fig:example-meta-data}
\end{figure*}

@@

\subsection*{Markers}
%
We make use of the full body 47 marker set described in \cite{Ton's HBM paper}
and presented in detail in Table \ref{tab:marker-labels}. As with all camera
based motion capture systems, the markers sometimes go missing in the
recording.

@@

\begin{table*}

@@

\end{table*}

\section*{Perturbation Signals}
%
The data set uses a testing protocol containing a segment of normal walking,
followed by a series of pseudo-random longitudinal perturbations, and ending
with a second segment of normal walking. Three pseudo-random speed control
signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
generated through MATLAB and Simulink. Initially, a  random acceleration signal
was created using discrete-time Gaussian white noise and saturated at the maximum
belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt speed, and 
high-pass filtered (second-order Butterworth) to eliminate velocity drift. Mean 
velocity was then added to the signal and limited between 0 and 3.6 m/s. The 
cutoff frequencies of the high-pass filter, as well as the variance in the acceleration 
signal, were adjusted until the appropriate standard deviations for each mean 
velocity were obtained.

@@

\begin{figure*}
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt velocity input signals (purple) and recorded output
    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
    respectively}
  \label{fig:input_output}
\end{figure*}

@@

\begin{figure*}
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency analysis of the treadmill belt velocity input signal
                 (purple) and the recorded output velocity (blue)
                for average belt speeds of 0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure*}

\section*{Processed Data}

@@

\begin{figure*}
  \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
  \caption{Mean (solid, dashed) and $2\sigma$ (shaded) ($N=120$) joint angles
    and torques from both unperturbed (purple) and perturbed (blue) gait cycles
    from trial 20.}
  \label{fig:angle-torque-comparison}
\end{figure*}

@@

\begin{figure*}
  \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
  \caption{Box plots of the average belt speed, stride frequency, and stride
    length which compare 120 gait cycles for the unperturbed (purple) and
    perturbed (blue).}
  \label{fig:gait-cycle-stats-comparison}
\end{figure*}

\section*{Conclusion}

@@

\subsection*{Author contributions}
A.v.d.B. conceived of the experiments and protocol. J.K.M and S.K.H developed
the protocol, ran the experiments, and collected the data. J.K.M was the
primary author of the paper with significant contributions from S.K.H and
A.v.d.B. All authors were involved in the revision of the draft manuscript and
have agreed to the final content.

\subsection*{Competing interests}
The authors have no  financial, personal, or professional competing interests
that could be construed to unduly influence the content of this article.

\subsection*{Grant Information}
The work was partially funded by the State of Ohio Third Frontier Commission
through the Wright Center for Sensor Systems Engineering (WCSSE).

\subsection*{Acknowledgments}
We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.

%{\small\bibliographystyle{unsrt}
%\bibliography{bibfile}}
